Constraining the Interiors of Asteroids Through Close Encounters

### 2022-05-17 paper/appendices.tex
Fit less to test degeneracy
Updated paper

In section \ref{sec:scan-cadence}, we noted that posterior uncertainty as a function of observation cadence $\Delta t$ appears to increase suddenly near $\Delta t \sim T_\text{cad}=30-40$ min. In the left panel of figure \ref{fig:cad-contour}, we display contour plots of posterior uncertainty $\sigma(K_{\ell m})$ of the fit parameters as a function of both cadence $\Delta t$ and $P_\omega$ (left panel) or the relative orbit speed $t_\text{spin} / t_\text{orbit}$ (right panel). Both panels show the sudden increase in posterior uncertainty we named $T_\text{cad}$. Now $T_\text{cad}$ is visible as a function of frequency and relative orbit speed. In all cases, the value of $\gamma_0$ was set so that all data points achieve the same value of $\gamma$ at perigee.

\begin{figure*}
  \centering
  \includegraphics[width=0.49\textwidth]{figs/cad-period.pdf}\hfill
  \includegraphics[width=0.49\textwidth]{figs/cad-speed.pdf}
  \caption{Contour plots showing posterior uncertainties in the first- and second-order fit parameters as a function of cadence $\Delta t$ and dynamical time scales for the encounter: rotational period $P_\omega$ (\textit{left}) and the relative speed of the orbit (\textit{right}; see text for a definition). The reference values of $P_\omega=9$ hr and $t_\text{spin}/t_\text{orbit}=1$ are shown as dotted lines. The red line represents the $\sigma(K_{\ell m})$ threshold. Cadence cutoff depends strongly on both $P_\text{omega}$ and $t_\text{spin}/t_\text{orbit}=1$.}
  \label{fig:cad-contour}
\end{figure*}

The relative orbit speed $t_\text{spin} / t_\text{orbit}$ was determined by un-physically increasing or decreasing the time at which the asteroid moved through the orbit determined by the equations of motion (but leaving the orbit shape unchanged). The equations of motion affecting the orientation and spin of the asteroid however were unaffected. With this unphysical process, we isolate the effect of the amount of time spent near perigee on posterior uncertainty, without inheriting additional affects that would have been caused by the orbit changing shape.

Figure \ref{fig:cad-contour} demonstrates that, near the reference values of $P_\text{omega}$ and $t_\text{spin} / t_\text{orbit}$, $T_\text{cad}$ depends on both time scales $P_\omega$ and $t_\text{spin} / t_\text{orbit}$. Specifically, high rotational period and slow orbits are optimal for large $T_\text{cad}$. The dependence on $P_\omega$ agrees with the fact that large rotational periods for fixed cadence lead to better posterior uncertainty, discussed in section \ref{sec:scan-period}. The figure also demonstrates that for large $P_\omega$, $T_\text{cad}$ depends less strongly on $P_\text{omega}$ and may even reverse its dependence such that increasing $P_\text{omega}$ decreases $T_\text{omega}$. In all cases except $\Re K_{33}$, at least constant-$\sigma(K_{\ell m})$ contour is seen to curve back such that $\Delta t$ decreases as a function of $P_\text{omega}$ for large $P_\text{omega}$. However, in these regions the cadence cut-off also dulls, as shown by the spreading of the constant-$\sigma(K_{\ell m})$ contours in this region.

The right panel of figure \ref{fig:cad-contour} shows a slightly stronger and more monotonic dependence of $T_\text{cad}$ on $t_\text{spin} / t_\text{orbit}$. It appears that slightly slower orbits sharply increases $T_\text{cad}$. Similar in the $P_\omega$ case, for increasingly fast orbits, $T_\text{cad}$ depends less on $t_\text{spin}/ t_\text{orbit}$. 
\jtd{I wanted to talk to Julien about the interpretation of this panel. Because $t_\text{spin} / t_\text{orbit}$, I find it hard to interpret. Maybe a different type of scan is better.}

Figure \ref{fig:cad-contour} demonstrates that, for large period, the cadence cut-off $T_\text{cad}$ reduces in prominence and rises to above one hour, meaning that large gaps of time between observations can still lead to well-constrained parameters in these cases. Faster orbits also lead to $T_\text{cad} > 1$ hr for for as small a decrease as $t_\text{spin} / t_\text{orbit}  \approx 75$\%.

>>>>

In section \ref{sec:scan-cadence}, we noted that posterior uncertainty as a function of observation cadence $\Delta t$ appears to increase suddenly near $\Delta t \sim T_\text{cad}=30-40$ min. In this appendix, we discuss the location of this cadence cut-off as a function of the physical parameters of the asteroid.

In figure \ref{fig:cad-contour}, we display contour plots of posterior uncertainty $\sigma(K_{\ell m})$ of the fit parameters as a function of both cadence $\Delta t$ and $P_\omega$ (left panel) or the relative orbit speed $t_\text{spin} / t_\text{orbit}$ (right panel). Both panels show the same sudden increase in posterior uncertainty we named $T_\text{cad}$, now visible as a function of frequency and relative orbit speed. In all cases, the value of $\gamma_0$ was set so that all data points achieve the same value of $\gamma$ at perigee.

\begin{figure*}
  \centering
  \includegraphics[width=0.48\textwidth]{figs/cad-period.pdf}\hfill
  \includegraphics[width=0.48\textwidth]{figs/cad-speed.pdf}
  \caption{Contour plots showing posterior uncertainties as a function of cadence $\Delta t$ and dynamical time scales for the encounter: rotational period $P_\omega$ (\textit{left}) and the relative speed of the orbit (\textit{right}; see text for a definition). The reference values of $P_\omega=9$ hr and $t_\text{spin}/t_\text{orbit}=1$ are shown as dotted lines. The red line represents the uncertainty threshold. Cadence cutoff depends strongly on both $P_\text{omega}$ and $t_\text{spin}/t_\text{orbit}=1$.}
  \label{fig:cad-contour}
\end{figure*}

Figure \ref{fig:cad-contour} demonstrates that large rotational period produces high $T_\text{cad}$. The dependence on $P_\omega$ agrees with the fact that large rotational periods for fixed cadence lead to better posterior uncertainty, discussed in section \ref{sec:scan-period}. The figure also demonstrates that for large $P_\omega$, $T_\text{cad}$ depends less strongly on $P_\text{omega}$ and may even reverse its dependence such that increasing $P_\text{omega}$ decreases $T_\text{omega}$. In all cases except $\Re K_{33}$, at least constant-$\sigma(K_{\ell m})$ contour is seen to curve back such that $\Delta t$ decreases as a function of $P_\text{omega}$ for large $P_\text{omega}$. In these regions the cadence cut-off also dulls, as shown by the spreading of the constant-$\sigma(K_{\ell m})$ contours in this region.

 The relative orbit speed $t_\text{spin} / t_\text{orbit}$ was defined by un-physically increasing or decreasing the time at which the asteroid moved through the orbit determined by the equations of motion (but leaving the orbit shape unchanged). The equations of motion affecting the orientation and spin of the asteroid however were unaffected. $t_\text{spin} / t_\text{orbit} > 1$ corresponds to a faster orbit, and $t_\text{spin} / t_\text{orbit} < 1$ corresponds to a slower orbit. With this unphysical process, we isolate the effect of the amount of time spent near perigee on posterior uncertainty, without inheriting additional affects that would have been caused by the orbit changing shape.

The right panel of figure \ref{fig:cad-contour} shows a stronger and more monotonic dependence of $T_\text{cad}$ on $t_\text{spin} / t_\text{orbit}$. It appears that even slightly slower orbits sharply increases $T_\text{cad}$. This effect is both due to the asteroid spending greater time in the high-torque, near-perigee region, and the larger data set that can be collected for slow orbits. However, if the orbit speed is changed by adjusting its parameters ($v_\infty$, $r_p$) or the central body mass $\mu_\mathcal{B}$, then the orbit shape will also change. This induces other effects studied in the main text and will complicate the trend observed here. Unlike the $P_\omega$ case, the cadence cut-off does not visibly broaden as a function of $t_\text{spin} / t_\text{orbit}$. It appears that the increased orbit speed merely shifts $T_\text{cad}$ rather than changing its sharpness.

### 2022-05-17 paper/main.tex
Updated paper

It will be useful to have a rule-of-thumb ``posterior uncertainty threshold,'' above which we say that posterior uncertainty $\sigma(K_{\ell m})$ is too large to yield meaningful estimates of density distribution. There are many possible definitions of posterior uncertainty, but we explore one in this paper which is outlined in appendix \ref{app:sigma-threshold}. This leads to cut-offs of the encounter properties which lead to posterior uncertainty less than the threshold. These critical values are shown in table \ref{tab:threshold-summary} as a summary, and are discussed in detail in the following sections.

\begin{table}
  \centering
  \begin{tabular}{lll} \hline
    Encounter property & Strong & Weak \\ \hline
    Excess velocity ($v_\infty$) & & \\
    Perigee ($r_p$) & & \\
    Spin pole uncertainty ($\sigma_\theta$) & & \\
    Spin period uncertainty ($P_\omega \sigma_\rho$) & & \\
    Asteroid length ($a_\mathcal{A}$) & & \\
    Cadence ($\Delta t$) & & \\
    Data gap ($T_\text{gap}$) & & \\
    Rotational period ($P_\omega$) & & \\ \hline
  \end{tabular}
  \caption{For each encounter property discussed in section \ref{sec:discussion}, the cut-offs (i.e., the values which leads to $\sigma(K_{3 m})$ at the threshold value) are given. Properties inside the strong cut-off have all $\sigma(K_{3m})$ under the threshold value, while properties inside the weak cutoff have at least one $\sigma(K_{3m})$ under the threshold value. The reference asymmetric asteroid is used. \jtd{summary}}
  \label{tab:threshold-summary}
\end{table}

>>>>

\jtd{It will be useful to have a rule-of-thumb ``posterior uncertainty threshold,'' above which we say that posterior uncertainty $\sigma(K_{\ell m})$ is too large to yield meaningful estimates of density distribution. We define this posterior uncertainty threshold as the point at which the median of the ratio between the the local density uncertainty and local density  ($\rho_\sigma / \rho$) is equal to a fraction $f$. For this analysis, we investigate both $f=100\%$ (corresponding to an essentially unconstrained density distribution) and $f=20\%$.}

\jtd{We use this threshold because it has clear physical meaning, unlike the density moments. But the value of $\rho_\sigma / \rho$ depends on the model used. In this analysis, we use the likelihood model, but choosing a different model with constraints added would reduce density uncertainty, and lead to less restrictive thresholds.}

\begin{table}
  \centering
  \begin{tabular}{lll} \hline
    Encounter property & $\sigma_\rho / \rho = 100\%$ & $20\%$ \\ \hline
    Perigee ($r_p$, Earth radii) & $<$7.9 & $<$4.7 \\
    Spin pole uncertainty ($\sigma_\theta$, $^\circ$) & $<$2.8 & $<$0.5 \\
    Spin period uncertainty ($P_\omega \sigma_\rho$, ms) & $<$15 & $<$3.1\\
    Asteroid length ($a_\mathcal{A}$, m) & $>$180 & $>$1100 \\
    Cadence ($\Delta t$, min) & $<$38 & $<$2 \\
    Data gap ($T_\text{gap}$, hr) & $<$1.8 & $<$0 \\
    Period ($P_\omega$, hr) & $>$4.3 & $>$9.5\\ \hline
  \end{tabular}
  \caption{For each encounter property discussed in section \ref{sec:discussion}, the properties that lead to density uncertainty equal to 100\% (20\%) of the local density are shown in the middle (right) columns. The reference asymmetric asteroid is used. The 20\% threshold is very restrictive, but the 100\% threshold is more achievable and demonstrates the thresholds in the right column are not hard cut-offs.}
  \label{tab:threshold-summary}
\end{table}
